\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-05 pldi/draft.tex
Improve Unix50 section

    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/topn_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:topn}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/wf_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:wf}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/shortest_scripts_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:spell}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
      \centering
      \includegraphics[width=\textwidth]{\detokenize{./figs/alt_bigrams_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:minimal_sort}
    \end{subfigure}%
    \caption{
      \textbf{Speedup achieved by \sys as a function of the level of parallelism (1--100).} 
      Three different configurations per benchmark:
      (i) parallel, focusing only on the parallel execution,
      (ii) +compile, adding \sys' pre-processing overheads, and
      (iii) \kk{Add with , without eager and split here}.
      \nv{Candidate values, based on the HW we have 2, 4, 8, 16, 32, 64. Also, maybe it makes sense to show log x-axis?}
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\heading{Programs}
In a recent event celebrating \unix's 50-year legacy, Bell Labs created a set of \todo{37} challenges~\cite{unix50} solvable by means of composing \unix pipelines.
By scanning GitHub, we were able to find a set of shell programs that solve all-but-one problems~\cite{}.
We consider this a good set of benchmarks because 
(i) the problems were designed to to highlight the \unix philosophy~\cite{} (rather than having a single command that solves the problem) and make extensive use of \unix built-ins with a plethora of flags, and (ii) the solutions were written by non-experts (contrary to the \S\ref{ours}) and were executed by \sys unmodified---stressing patterns that we, \sys's authors, did not necessarily anticipate.
In cases where an obvious fix improves the pipeline's performance, we report on both \sys's original speedup and the one achieved by applying the fix.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

@@

\TODO{An important point that needs to be made clear is that these are
  taken from the wild unmodified. And that Pash never leads to
  slowdown (except for the trivial head cases, where the slowdown is
  neligible).}

\heading{Results}
\sys's parallelism set to \todo{16}. The input given to the pipelines
was the test input from unix50 multiplied to reach \todo{10GBs}. The
speedup of each pipeline separately is shown in
\Cref{fig:unix50-individual}. The average\todo{/median} speedup is
\todo{$5.24\times$} and the \todo{geometric mean is $XX\times$}.

@@

\heading{Take-aways}
\sys leads to significant performance improvements in completely unmodified \unix pipelines found out in the wild.
Small tweaks can yield further improvements, showing that \sys-awareness and scripting expertise can improve results.

@@

\kk{Out of all benchmarks, we have talked about the head and the
  non-parallelizable ones in the paper. Maybe we also want to talk
  about the long and the cheap ones. Also we might want to talk and
  find out more about the sort bottleneck.}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Speedup (left axis) and over sequential execution (right axis) for all Unix50 pipelines.}
		Parallelism is \todo{16$\times$} on \todo{10GB} of input data~\cf{unix50}.
		\nv{Can decrease height and increase width substantially?}
    %% \todo{Scale the 0-1 region of the yaxis as much as the 1-20.}
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

\sys achieves speedup in all pipelines except for 2, 19, and 31. These
pipelines include a \ttt{head} and only process one line of the
input. This can be seen by their sequential execution time which is in
the order of 10 milliseconds. The slowdown is due to constant costs of
setting up the pipes and spawning many processes. In all three of
these cases the execution time with \sys is in the order of \todo{X
  ms}. For pipelines 13, 24, 25, 26, 29, and 30, \sys performs the
same as the sequential implementation. This is because these pipelines
contain non-parallelizable commands, such as \ttt{awk}, \ttt{sed} with
option \ttt{d}, and \ttt{tr} with option \ttt{-d}. An illustrative
example is pipeline 13, which is shown here:

@@

By executing this version of the pipeline with \sys, we achieve
\todo{X} speedup. This indicates that just a little awaresess of
command parallelizability can go a long way, and that the effort that
a \sys user needs to spend is often negligible to achieve
benefits. Finally, an important point that is reinforced from this
experiment is that \sys doesn't lead to slowdown, even in cases where
parallelism is not possible.

@@

>>>>

    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/topn_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:topn}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/wf_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:wf}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
        \centering
        \includegraphics[width=\textwidth]{\detokenize{./figs/shortest_scripts_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:spell}
    \end{subfigure}%
    ~
    \begin{subfigure}[b]{0.24\textwidth}
      \centering
      \includegraphics[width=\textwidth]{\detokenize{./figs/alt_bigrams_throughput_scaleup.pdf}}
        %% \caption{TODO}
        %% \label{eval:minimal_sort}
    \end{subfigure}%
    \caption{
      \textbf{Speedup achieved by \sys as a function of the level of parallelism (1--100).} 
      Four different configurations per benchmark:
      (i) \todo{Parallel}, showing the performance of \sys with both eager and split optimizations enabled,
      (ii) \todo{No Split}, showing the performance of \sys with split disabled,
      (iii) \todo{No Eager}, showing the performance of \sys with split and eager disabled,
      (iv) \todo{Blocking Eager}, showing the performance of \sys with split disabled and the naive blocking eager enabled,
      \nv{Maybe it makes sense to show log x-axis?}
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\heading{Programs} In a recent event celebrating \unix's 50-year
legacy, Bell Labs created a set of \todo{36} challenges~\cite{unix50}
solvable by means of composing \unix pipelines.  By scanning GitHub,
we were able to find a set of shell programs that solve
\todo{all-but-three} problems~\cite{}.  We consider this a good set of
benchmarks because (i) the problems were designed to to highlight the
\unix philosophy~\cite{} (rather than having a single command that
solves the problem) and make extensive use of \unix built-ins with a
plethora of flags, and (ii) the solutions were written by non-experts
(contrary to the \S\ref{ours}) and were executed by \sys
unmodified---stressing patterns that we, \sys's authors, did not
necessarily anticipate \kk{I would change that to ``stressing a
  variety of patterns'' since it doesn't really matter if we
  anticipated them (especially since we can address any issues and
  then re-run them).} .  In cases where an obvious fix improves the
pipeline's performance, we report on both \sys's original speedup and
the one achieved by applying the fix.
% TODO: talk about correct wrt developer intentions vs. correct wrt problem definition

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}

\heading{Results} \sys's parallelism set to 16. The input given to the
pipelines was the test input from unix50 multiplied to reach
10GBs. The speedup of each pipeline separately is shown in
\Cref{fig:unix50-individual}. The average\todo{/median} speedup is
$5.49$/\todo{$6.07\times$}, the \todo{geometric mean is $1.08\times$},
and the \todo{weighted average (with the absolute times as weights) is
  $5.75\times$}.

%% Mean: 5.499164686559543
%% Median: 6.075774436154179
%% Geometric Mean: 1.0798740409683736
%% Weighted Average: 5.752406752503295

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Speedup (left axis) and over sequential execution (right axis) for all Unix50 pipelines.}
		Parallelism is 16$\times$ on 10GB of input data~\cf{unix50}.
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

@@

@@

%% \kk{Out of all benchmarks, we have talked about the head and the
%%   non-parallelizable ones in the paper. Maybe we also want to talk
%%   about the long and the cheap ones. Also we might want to talk and
%%   find out more about the sort bottleneck.}

@@

\sys achieves speedup in all pipelines except for 2, 19, and 31. These
pipelines include a \ttt{head} and only process one line of the
input. This can be seen by their sequential execution time which is in
the order of 10 milliseconds. The slowdown is due to constant costs of
setting up the pipes and spawning processes. In all three of these
cases the execution time with \sys is in the order of 1 second. For
pipelines 13, 24, 25, 26, 29, and 30, \sys performs the same as the
sequential implementation. This is because these pipelines contain
commands that can not be parallelized in general, such as \ttt{awk},
\ttt{sed} with option \ttt{d}, and \ttt{tr} with option \ttt{-d}. An
illustrative example is pipeline 13, which is shown here:

@@

By executing this version of the pipeline with \sys, we achieve
\todo{X} speedup. This indicates that just a little awaresess of
command parallelizability can go a long way, and that the effort that
a \sys user needs to spend is often negligible to achieve
benefits.

Finally, for the rest of the pipelines \sys speedup is \todo{capped}
because of a combination of the following reasons: (i) they contain
pure commands that are parallelizable but don't scale linearly, such
as \ttt{sort} (e.g. pipelines 0, 1, 3, 15, 16, 18, 20, 27, 28, 33),
(ii) they are deep pipelines that already exploit task-based
parallelism (e.g. pipelines 7, 8, 9, 10, 11, 12, 14, 27, 28), and
(iii) they are not cpu-intensive, leading to \todo{larger IO and
  constant costs} (e.g. pipelines 4, 5, 6, 7, 8, 12, 14, 22, 23).

%% Sort-bottleneck: 1, 3, 15, 16, 20, 33
%% Non-parallelizable completely: 13 (awk-solved), 18 (awk), 24, 25, 26, 29, 30
%% Cheap (non cpu intensive -- constant costs grow): 0, 4, 5, 6, 7, 8, 12, 14, 22, 23
%% Long (task-based parallelism in sequential and context switches): 7, 8, 9, 10, 11, 12, 14, 21, 27, 28

%% Finally, an important point that is reinforced from this
%% experiment is that \sys doesn't lead to slowdown, even in cases where
%% parallelism is not possible.

\heading{Take-aways} \sys leads to significant performance
improvements in completely unmodified \unix pipelines found out in the
wild. Small tweaks can yield further improvements, showing that
\sys-awareness and scripting expertise can improve
results. Furthermore, it doesn't cause slowdown in any non-trivial
computation, even in cases where parallelism is not possible.

@@

\kk{Maybe use phoenix for that one? Since it has a MapReduce interface
  but it runs on multicore?}
